---
title: "TheFinalProject"
date: "2023-04-26"
output:
  word_document: default
---


### Import libraries and connect to SQL
```{r}
library(tidyverse)
library(tidymodels)
library(dplyr)
library(DBI)
library(odbc)
library(fastDummies)
library(boot)
library(plyr)
library(neuralnet)
library(caret)
library(gridExtra)
library(reshape2)
library(TTR)
library(randomForest)


# conn <- DBI::dbConnect(
#   odbc::odbc(),
#   Driver="SQL Server",
#   Server="LAPTOP-3AMVCH9J\\SQLEXPRESS",
#   Database="QuantData",
#   options(connectionObserver = NULL)
# )

norm = function(x){
  m0 = min(x)
  m1 = max(x)
  result = (x - m0)/(m1 - m0)
  return(result)
}

regular = function(x, y){
  m0 = min(y)
  m1 = max(y)
  result = x*(m1 - m0) + m0
  return(result)
}


```
# Machine Learning Round One
### Fetch datasets from SQL
```{r}
# bike_donations <- dbGetQuery(conn, "SELECT TOP 50000 * FROM dbo.BikeDonations")
# bike_events <- dbGetQuery(conn, "SELECT * FROM dbo.BikeEvents")
#write_csv(bike_donations, "BikeDonations.csv")
#write_csv(bike_events, "BikeEvents.csv")
```
### Read generated csv files
```{r}
bike_donations <- read_csv("BikeDonations.csv")
bike_events <- read_csv("BikeEvents.csv")
```
### Join the tables and omit N/A variables
```{r}
df2 <- left_join(x=bike_donations, y=bike_events, by="EventID")
df2[df2 == "N/A"] <- NA
df2 <- df2 %>% na.omit(df2)
```
### Convert string variables from dataset into numeric
```{r}
df3 <- df2 %>%
  mutate(GiftAmount=as.numeric(GiftAmt.x),
         Goals=as.numeric(Goals),
         ActiveReg=as.numeric(ActiveReg),
         NoReg=as.numeric(TotalFees),
         SentEmails=as.numeric(SentEmails)) %>%
  select(-EventID,-FiscalYear.x,-GiftAmt.x,-GiftAmt.y,-TotalFees,-ConfirmedGifts,-TotalOnlineGifts,-FiscalYear.y,-CampID,-DonorConsID,-Goals,-TeamID)

```
### Fix some of the variables spacing and such
```{r}
df3[df3 == "I have a Friend or Co-worker with MS"] <- "FriendOrCoWorker"
df3[df3 == "Bad (Soft Bounce)"] <- "SoftBounce"
df3[df3 == "Bad (Hard Bounce)"] <- "HardBounce"
df3[df3 == "Relative: Parent of person with MS"] <- "RelativeParent"
df3[df3 == "Relative: Other"] <- "RelativeOther"
df3[df3 == "I have a Friend of Co-worker with MS"] <- "FriendOfCoWorker"
```
### Parse dummy variables in dataset
```{r}
dataset <- fastDummies::dummy_cols(df3) %>%
  select(-GiftType,-PmtMethod,-Registered,-EmailStatus,-Connection)
colnames(dataset) = gsub(" ", "_", colnames(dataset))


pre_norm_set <- dataset
```
### Extract Column Names
```{r}
colnames(dataset)
```
### Calculate correlation matrix of variables
```{r}
library(corrplot)
corrplot(cor(dataset), method = 'ellipse', order = 'AOE', type = 'upper', tl.cex = 0.5)
```
### Prepare data for machine learning: Round One
```{r}
set.seed(1337)
data_split <- initial_split(pre_norm_set, prop=0.7)
data_train <- data_split %>% training()
data_test <- data_split %>% testing()

dataset <- pre_norm_set %>%
  mutate(
    GiftAmount=norm(GiftAmount),
    ActiveReg=norm(ActiveReg),
    NoReg=norm(NoReg),
    SentEmails=norm(SentEmails)
  )

norm_split <- initial_split(dataset, prop=0.7)
norm_train <- norm_split %>% training()
norm_test <- norm_split %>% testing()
```
### Model A1: Neural Network
```{r}
nnA <- neuralnet(GiftAmount ~ ., data=norm_train, hidden=c(8, 4))
plot(nnA)
```
### Calculate RMSE for the first runs neural net model
```{r}
predA <- compute(nnA, norm_test)
predA <- regular(predA$net.result, data_test$GiftAmount)
xx1 <- data_test$GiftAmount

RMSE_NN_ModelA <- (sum((xx1 - predA)^2) / length(xx1)) ^ 0.5
cat('RMSE for Neural Network Model A1: ', RMSE_NN_ModelA)
```
### Graph the actual versus the predicted values
```{r}
ggplot(mapping=aes(x=xx1, y=predA)) +
  geom_point() +
  labs(title="NN Model A", x="Actual", y="Predicted")
```
### Create a log version of the plot above
```{r}
options(scipen=999)
ggplot(mapping=aes(x=xx1, y=predA)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(title="NN Model A (Log)", x="Actual", y="Predicted")
```
### Model A2: MultiVariable Regression
```{r}
reg_modelB1 <- lm(GiftAmount ~ ., data=data_train)
summary(reg_modelB1)
```
### Compute the RMSE of the Regression model
```{r}
predB1 <- reg_modelB1 %>% predict(data_test)

rt <- data_test$GiftAmount

RMSE_RegModelB1 <- (sum((rt - predB1)^2)/length(rt))^0.5
cat('RMSE for Regression Model B1: ', RMSE_RegModelB1)
```
### Graph the results of the first regression model
```{r}
ggplot(mapping=aes(x=rt, y=predB1)) +
  geom_point() +
  labs(title="LinReg Model B1", x="Actual", y="Predicted")
```
### Graph the previous graph using the log scale
```{r}
ggplot(mapping=aes(x=rt, y=predB1)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(title="LinReg Model B1 (Log)", x="Actual", y="Predicted")
```
### Print all significant variables
```{r}
coef1 <- data.frame(summary(reg_modelB1)$coef[summary(reg_modelB1)$coef[,4] <= .05, 4])
coef1
```
# Machine Learning Round Two
### Run your second neural network model
```{r}
nnB <- neuralnet(GiftAmount ~ GiftType_offline + PmtMethod_Cash + EmailStatus_Good + EmailStatus_HardBounce + EmailStatus_SoftBounce + Connection_Blank + Connection_Caregiver_of_Person_with_MS + Connection_Child_has_MS + Connection_Friend_has_MS + Connection_I_have_MS + Connection_None + Connection_Other + Connection_Parent_has_MS + Connection_Possible_MS + Connection_Relative_has_MS + Connection_Sibling_has_MS, data=norm_train, hidden=c(8, 4))
plot(nnB)
```
### Calculate the RMSE for the second neural network
```{r}
predA2 <- compute(nnB, norm_test)
predA2 <- regular(predA2$net.result, data_test$GiftAmount)
xx12 <- data_test$GiftAmount

RMSE_NN_ModelA2 <- (sum((xx12 - predA2)^2) / length(xx12)) ^ 0.5
cat('RMSE for Neural Network Model A2: ', RMSE_NN_ModelA2)
```
### Graph the actual and expected variables based off the second neural network model
```{r}
ggplot(mapping=aes(x=xx12, y=predA2)) + 
  geom_point() +
  labs(title="Neural Net Prediction Model 2", x="Actual", y="Predicted")
```
### Log plot of neural network model 2
```{r}
ggplot(mapping=aes(x=xx12, y=predA2)) + 
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(title="Neural Net Prediction Model 2 (Log)", x="Actual", y="Predicted")
```
### Calculate your second linear regression with the significant variables
```{r}
reg_modelB2 <- lm(GiftAmount ~ GiftType_offline + PmtMethod_Cash + EmailStatus_Good + EmailStatus_HardBounce + EmailStatus_SoftBounce + Connection_Blank + Connection_Caregiver_of_Person_with_MS + Connection_Child_has_MS + Connection_Friend_has_MS + Connection_I_have_MS + Connection_None + Connection_Other + Connection_Parent_has_MS + Connection_Possible_MS + Connection_Relative_has_MS + Connection_Sibling_has_MS, data=data_train)
summary(reg_modelB2)
```
### Calculate RMSE off you second regression model
```{r}
predB2 <- reg_modelB2 %>% predict(data_test)

rt2 <- data_test$GiftAmount

RMSE_RegModelB2 <- (sum((rt2 - predB2)^2)/length(rt2))^0.5
cat('RMSE for Regression Model B2: ', RMSE_RegModelB2)
```
### Plot the actual vs. predicted values for your second regression
```{r}
ggplot(mapping=aes(x=rt2, y=predB2)) +
  geom_point() +
  labs(title="Regression Model 2 Actual vs. Predicted", x="Actual", y="Predicted")
```
### Give the log plot of the plot above
```{r}
ggplot(mapping=aes(x=rt2, y=predB2)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(title="Regression Model 2 Actual vs. Predicted (Log)", x="Actual", y="Predicted")
```
# Ensemble Method: Combining Neural Network With Decison Tree
```{r}
ensemble_set <- pre_norm_set %>%
  select(ActiveReg, NoReg, SentEmails,GiftAmount, EmailStatus_Good, Connection_Friend_has_MS)

set.seed(567)
ensemble_split <- initial_split(ensemble_set, prop=0.7)
ensemble_train <- ensemble_split %>% training()
ensemble_test <- ensemble_split %>% testing()

ensemble_set <- ensemble_set %>%
  mutate(
    ActiveReg=norm(ActiveReg),
    NoReg=norm(NoReg),
    SentEmails=norm(SentEmails),
    GiftAmount=norm(GiftAmount)
  )

norm_e_split <- initial_split(ensemble_set, prop=0.7)
norm_e_train <- norm_e_split %>% training()
norm_e_test <- norm_e_split %>% testing()

```
### Build a neural network model
```{r}
nnC <- neuralnet(GiftAmount ~ ., data=norm_e_train, hidden=c(5, 3))
plot(nnC)
```
### Alter the dataset to hold the predictions
```{r}
predictions <- compute(nnC, norm_e_test)
netResults <- regular(predictions$net.result, ensemble_test$GiftAmount)

# Replace the actual amount with the predicted amount, bin Gift Amount

dset <- ensemble_test %>%
  select(-GiftAmount) %>%
  mutate(GiftAmount=cut(netResults, breaks=c(-1, 90, 150), labels=c("Low Donation", "High Donation")))

table(unlist(dset[, c("GiftAmount")]))
```
### Plot Histogram of Gift Amount Class
```{r}
dset %>%
  ggplot(mapping=aes(x=GiftAmount)) +
  geom_histogram(stat="count")
```

### Split new dataset
```{r}
#set.seed(999)
tree_split <- initial_split(dset, prop=0.7)
tree_train <- tree_split %>% training()
tree_test <- tree_split %>% testing()


```
### Train decision tree
```{r}
library(rpart)
library(rpart.plot, warn.conflicts=FALSE)

fit <- rpart(GiftAmount ~ ., data=tree_train, method="class",control=rpart.control(cp=0))

rpart.plot(fit, extra=100)
```
### Generate Classifications from the Decision Tree (Confusion Matrix)
```{r}
prediction <- predict(fit, tree_test, type='class')
confusionMatrix(prediction, tree_test$GiftAmount, mode="everything")
```
### Compute ROC Curve and AUC
```{r}
library(pROC)
prob_pred <- predict(fit, tree_test, type='prob')[,2]

roc_curve <- roc(tree_test$GiftAmount, prob_pred)
plot(roc_curve, main="ROC Curve", print.auc=TRUE, legacy.axes=TRUE, revC=TRUE)
auc(roc_curve)
```
```{r}

```
```{r}

```
```{r}

```
```{r}

```